---
title: "HW6"
author: "Dimitrios Mitras"
date: "2024-01-26"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


In this project, an analysis is performed on the dataframe = "album". The 'album' dataframe consists of the genre, artist, and album information (year and title). Each album contains multiple genres. 
For easier data management, an 'album_all' dataframe was created using Excel Power Query, where each genre is placed in a separate column.

<h2> Import libraries and datasets </h2>
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
library(MASS)
library(dplyr)
library(tidyr)
library(readxl)
```

```{r message=FALSE, warning=FALSE}
library(readr)
album <- read_csv("album.csv")

#genres.xlxs create by Excel queries
album_all <- read_excel(("genres2.xlsx"))

#dataframe with artist and his genres (problem 4)
#Delete duplicates of artists
genres <- album_all %>%
  dplyr::select(-album) %>%
  distinct(artist , .keep_all = TRUE)
```


```{r}
print(genres)
```





<h3> Data Preprocessing </h3>
```{r}
# Check Number of missing values in each column in the data frame
colSums(is.na(album))
```
```{r}
#genre.1 and artist is what we are certain of 
colSums(is.na(genres))
```

```{r}
#Separate year from album title
album <- album %>%
  mutate(album = as.character(album)) %>%
  separate(album , into = c("year" , "album"), sep= " " , extra = "merge")

#Remove parentheses from the 'year' column
album$year <- gsub("[()]", "", album$year)

#Separate year from album title
album_all <- album_all %>%
  mutate(album = as.character(album)) %>%
  separate(album , into = c("year" , "album"), sep= " " , extra = "merge")

#Remove parentheses from the 'year' column
album_all$year <- gsub("[()]", "", album_all$year)
```


```{r}
#Create dataframe artist 
artist<- unique(album[,c("artist" , "genre")])

#Small difference between length(artist) and length (genres)
#3694 vs 3692
#Probably sth with lower and uppercase
artist <- artist %>%
  distinct(artist , .keep_all = TRUE)

#Now its okay
```

Create data.frame artist_vector with all pairs (artist - genre)

```{r}
artist_vector <- artist %>%
  mutate(genre_list = strsplit(genre, ","))

artist_vector <- unnest(artist_vector, genre_list)

artist_vector <- artist_vector [,-2]
 

```





<h3> Percentage distribution of artists based on their primary genre </h3>
```{r}
#1a
#Count artists based on their 1st genre
genre_count_first <- genres %>%
  group_by(genre.1) %>%
  summarise(artist_count = n_distinct(artist))

# Calculate the percentage for each genre
genre_count_first <- genre_count_first %>%
  mutate(percentage = (artist_count / sum(artist_count)) * 100)

#create a theme_minimal which does not display legend in plots.
thm <- theme_minimal() +
  theme(
    legend.position = "none"
  ) 

#Plot genre_count_first
ggplot(genre_count_first, aes(x= genre.1 , y=percentage, fill = genre.1))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = sprintf("%.1f%%", percentage)), vjust = -0.5, size = 3)  +
  thm +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title= "Bar plot percentages of artists based on first genre")
```




<h3> Percentage distribution of artists based on all their genre </h3>
```{r}
#1b
#Count artists based on all their genres
genre_count_all<- artist_vector %>%
  group_by(genre_list) %>%
  summarise(artist_count = n_distinct(artist))

# Calculate the percentage for each genre
genre_count_all <- genre_count_all %>%
  mutate(percentage = (artist_count / sum(artist_count)) * 100)

#Plot genre_count_all
ggplot(genre_count_all, aes(x= genre_list , y=percentage, fill = genre_list))+ 
  geom_bar(stat = "identity")+
  geom_text(aes(label = sprintf("%.1f%%", percentage)), vjust = -0.5, size = 3)  +
  thm +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title= "Bar plot percentages of artists based on all genres")
```





<h3> Distribution of albums per year </h3>

```{r}
#2. Albums per year
albums_per_year<- album %>% dplyr::select(-artist, -genre ) 

#Create frame with count of albums per year
albums_per_year <- albums_per_year %>%
  group_by(year) %>%
  summarize(count = n_distinct(album)) 

#print list
knitr::kable(albums_per_year, format = "markdown")


```





<h3> Distributions of albums for the top 5 genres over time </h3>

Data frame Creation. Identification of the top 5 genres based on the primary genre of each album.
```{r message=FALSE, warning=FALSE}
#3
#Keep only first genre for each album
genre_first_album <- album_all%>%
  dplyr::select(genre.1 , year , album)

#For every genre count each albums, keep 5 first
top5_genres_names <- genre_first_album %>%
  group_by(genre.1)%>%
  summarize (count_genre = n_distinct(album)) %>%
  arrange(desc(count_genre)) %>%
  head(5) 

#Keep only genre names
top5_genres_names <- top5_genres_names %>%
  dplyr::select(genre.1)

#Compute for every genre total albums per year
genres_count_per_year <- genre_first_album %>%
  group_by(genre.1 , year)%>%
  summarize (count = n())

#Keep only the top5 from genres_count_per_year
#Join with top 5 genre names 
top5_genres_count_per_year <- genres_count_per_year %>%
  semi_join(top5_genres_names, by = "genre.1")
```
Line plotting of top5 genres' albums and their distribution per year
```{r}
#Plot for top 5 genres each albums count per year

#Convert column year to numeric
top5_genres_count_per_year$year <- as.numeric(top5_genres_count_per_year$year)

ggplot(top5_genres_count_per_year, aes (x=year, y=count,group = genre.1, color=genre.1))+
  geom_line() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(title= "Line plot of top5 genres' albums and their distribution per year")

```





<h3>Mosaic for primary and secondary genre </h3>

Below, the relationship between the primary and secondary genres is examined through a mosaic plot. In cases where there is no secondary genre, the same value as the primary genre has been filled in. Therefore, in the diagonal, the color is based on whether exists a unique genre.
```{r}
#4 

#From dataframe genre keep only artist, genre.1, genre.2
first_second_genres <- genres %>%
  dplyr::select(-genre.3 , -genre.4 , -genre.5)

#Add secondary_genre same as first if is NA
first_second_genres <- first_second_genres %>%
  mutate(secondary_genre = coalesce(genre.2 , genre.1))

#Rename genre.1 to first_genre
first_second_genres <- first_second_genres %>%
  rename(first_genre = genre.1)

#Delete column genre.2 and column artist
first_second_genres <- first_second_genres %>%
  dplyr::select(-genre.2 , -artist)
```

For each relationship, their frequency is calculated. Since the differences between the frequencies vary, their logarithms were calculated
```{r}
first_second_genres <- first_second_genres %>%
  group_by(first_genre , secondary_genre) %>%
  summarise(freq = log(n()))
```

Due to the vast number of different genres, it was preferred to display it using ggplot and not mosaic().
```{r}
#Plot mosaic with ggplot2
#log freq
ggplot(data = first_second_genres, aes(x = first_genre, y = secondary_genre , fill=freq)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "red") +
  labs(title = "Log Frequency between first and second genre") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```




<h3> Country names found in album titles </h3>

Import libraries for spatial data. For this analysis, spData::world data.frame will be used
```{r message=FALSE, warning=FALSE}
###5
#spatial data
library(sf)          
library(terra)     
library(spData) 

my_world <- spData::world

countries_names <- my_world$name_long
albums_names <- album_all$album 
 
```


Finding country names in albums through a loop structure.
```{r}
library(stringr)

#Initialize a vector to store the count for each country
country_counts <- integer(length(countries_names))

#For every country name detects if exist in album_names.
#Might an album_name contain more than one country.
for (i in seq_along(countries_names)) {
  country_counts[i] <- sum(str_detect(albums_names, countries_names[i]))
}

#Create a data frame with country names and their respective counts
result_df <- data.frame(country = countries_names, albums_exist = country_counts)

#Join result with dataset world
my_world <- left_join (my_world , result_df, by = c("name_long" = "country"))
```


Choropleth map with a color ramp and binning
```{r}

library(leaflet)

# Define the color palette
bins <- c(0, 1, 2, 3, 4, 5, 6, 7,22 )
colors <- colorBin("YlOrRd", my_world$albums_exist, bins=bins)

# Create a leaflet map
leaflet(my_world) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')
  )) %>%
  addPolygons(
    fillColor = ~colors(albums_exist),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = ~paste(name_long, ": ", albums_exist, " times"),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto"
    )
  ) %>%
  addLegend(pal = colors, values = ~albums_exist, opacity = 0.7, title = NULL,
            position = "bottomright")



```

